### 2015-08-26 ICRA2016/bandits_docs/paper.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8
\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
% fix acrobat png issue bug
\pdfminorversion=4
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\usepackage{balance}
%\usepackage{subfigure}
%\usepackage[subfigure]{tocloft}

%\usepackage[colorlinks]{hyperref}
%\usepackage[caption=false]{subfig}

\usepackage{caption}
%\usepackage{subcaption}
\captionsetup{compatibility=false}

\usepackage{benktools}
\usepackage{tikz}
\usetikzlibrary{external}
\usetikzlibrary{positioning,shapes,shadows,arrows, calc, fit, backgrounds, arrows, positioning}
\tikzexternalize[prefix=tikz/]

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\input{includes/preamble.tex}

\begin{document}
\title{\LARGE \bf DexNet 1.0: Correlated Multi-Armed Bandit Models for Large Scale 3D Grasp Planning Under Uncertainty
		\versionline{2}}
\author{Jeffrey Mahler, Brian Hou, Mel Roderick, Michael Laskey, Kai Kohlhoff, \\ Torsten Kroeger, Florian Pokorny, James Kuffner, Ken Goldberg}
%\institute{Deptartment of Electrial Engineering and Computer Sciences, University of California, Berkeley, USA
%\{jmahler, zmccarthy, sachinpatil, pabbeel, goldberg\}@berkeley.edu} 
\maketitle

\TODO{Possibly will include Matthieu Aubry and Pieter Abbeel as coauthors as well}

\input{includes/abstract}
%
\input{includes/introduction}
\input{includes/related_work}
%
\input{includes/problem_statement}
\input{includes/background}
\input{includes/algorithm}
%\input{includes/analysis}
\input{includes/experiments}
%\input{includes/conclusion}

%%\appendix
%\include{includes/proofs}

%\balance
\bibliographystyle{IEEEtranS}
\bibliography{bibliography}

\end{document}