### 2015-09-03 ICRA2016/bandits_docs/paper.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8
\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
% fix acrobat png issue bug
\pdfminorversion=4
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\usepackage{balance}
%\usepackage{subfigure}
%\usepackage[subfigure]{tocloft}

%\usepackage[colorlinks]{hyperref}
%\usepackage[caption=false]{subfig}

\usepackage{caption}
%\usepackage{subcaption}
\captionsetup{compatibility=false}

\usepackage{benktools}
\usepackage{tikz}
\usetikzlibrary{external}
\usetikzlibrary{positioning,shapes,shadows,arrows, calc, fit, backgrounds, arrows, positioning}
\tikzexternalize[prefix=tikz/]

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\input{includes/preamble.tex}

\begin{document}
\title{\LARGE \bf Dex-Net 1.0: A Cloud-Based Network of 3D Object Models Based on Deep Learning for Object Similarity and Multi-Armed Bandits for Robust Grasp Planning
%Correlated Multi-Armed Bandit Models for Large Scale 3D Grasp Planning Under Uncertainty
		\versionline{4}}
\author{Jeffrey Mahler$^1$, Florian Pokorny$^1$, Brian Hou$^1$, Mel Roderick$^1$, Michael Laskey$^1$, Kai Kohlhoff$^3$, \\ Torsten Kroeger$^3$, James Kuffner$^3$, Ken Goldberg$^2$
\thanks{$^1$Department of Electrical Engineering and Computer Sciences; {\tt\small \{jmahler, ftpokorny, brian.hou\}@berkeley.edu, melrose\_roderick@brown.edu}}%
\thanks{$^2$Department of Industrial Engineering and Operations Research; {\tt\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley, USA}%
\thanks{$^3$ Google Inc., Mountain View, USA; {\tt\small \{kohlhoff, tkr, kuffner\}@google.com}}
}
\maketitle

\TODO{Possibly will include Matthieu Aubry, Alyosha Efros, and Pieter Abbeel as coauthors} \\
\TODO{Update number of models in title: I want it to reflect exactly how many we use in experiments}

\input{includes/abstract}
%
\input{includes/introduction}
\input{includes/related_work}
%
\input{includes/problem_statement}
\input{includes/background}
\input{includes/algorithm}
\input{includes/dataset}
\input{includes/experiments}
\input{includes/conclusion}

%%\appendix
%\include{includes/proofs}

%\balance
\bibliographystyle{IEEEtranS}
\bibliography{bibliography}

\end{document}