### 2015-09-10 ICRA2016/bandits_docs/includes/algorithm.tex
Added many new figures as well as updated text. Quite a bit closer

\section{Correlated Multi-Armed Bandit Algorithm}
\seclabel{algorithm}
Our algorithm, detailed in Algorithm~\ref{alg:full}, plans a sequence of grasps with high $P_F$ for a new object $\mO$ using Multi-Armed Bandits (MABs) with correlated rewards and priors from the grasp and 3D object data in Dex-Net.
We first generate a set of candidate grasps $\Gamma$ for object $\mO$ using the antipodal grasp sampling described in \secref{grasp-sampling} and predict a prior distribution for each grasp using the Dex-Net database $\mD$.
Next, we run MAB by selecting the next grasp to sample using Thompson sampling~\cite{laskey2015bandits}, sampling from our uncertainty random variables, determining force closure for the grasp on the sampled variables as described in \secref{quality}, and updating a belief distribution on the probability of force closure for each grasp.
Finally, we return the grasp for which the belief distribution has the maximum lower confidence bound, a conservative estimate of the grasp with the highest estimated $P_F$.

\begin{algorithm}
{\small
    \SetAlgoLined
    {\bf Input:} Object $\mO$, Number of Candidate Grasps $N_g$, Number of Samples $N_s$, Number of Nearest Neighbors $N_n$, Dex-Net 1.0 Database $\mD$, Object Feature Map $\psi$,  Maximum Iterations $T$, Prior beta shape $\alpha_0$, $\beta_0$, Lower Bound Confidence $p$, Random Variables $\nu$, $\xi$, and $\gamma$ \\
    \KwResult{Estimate of the grasp with highest $P_F$, $\hat{\bg}^*$}

    \tcp{Generate candidate grasps and priors}
	$\Gamma$ = AntipodalGraspSample($\mO, N_g$) \;
	$\mA_0 = \varnothing, \mB_0 = \varnothing$\;
	\For{$\bg_k \in \Gamma$}{
		\tcp{Equations~\ref{eq:alpha-prior} and ~\ref{eq:beta-prior}}
		$\alpha_{k,0}, \beta_{k,0} =$ ComputePriors($\mO, \bg_k, \mD, N_n, \psi$)\; 
		$\mA_0 = \mA_0 \cup \{\alpha_{k,0}\}, \mB_0 = \mB_0 \cup \{\beta_{k,0}\}$\;
	}

	\tcp{Run MAB to Evaluate Grasps}
	\For{$t = 1, .., T$}{
		$j = $ ThompsonSample($\mA_{t-1}, \mB_{t-1}$)\;
		$\hat{\nu}, \hat{\xi}, \hat{\gamma} =$ SampleRandomVariables($\nu, \xi, \gamma)$\;
		$F_{j} = $ EvaluateForceClosure($\bg_j, \hat{\nu}, \hat{\xi}, \hat{\gamma}$)\;
		\tcp{Equations~\ref{eq:alpha} and ~\ref{eq:beta}}
		$\mA_{t}, \mB_{t} = $ UpdateBeta($j, F_j, \Gamma$)\; 
		$\bg_{t}^* = $MaxLowerConfidence($\mA_{t}, \mB_{t}, p$)\;
	}
	return $\bg_T^*$\;
    \caption{Robust Grasp Planning With Multi-Armed Bandits Using Dex-Net 1.0}
    \label{alg:full}
}
\end{algorithm}

\subsection{Predicting Grasp Quality Using Prior Data}
\seclabel{ccbps}
We use Continuous Correlated Beta Processes (CCBPs) to model correlations between the $P_F$ between grasps on different objects to utilize prior grasp and object data from Dex-Net 1.0.
CCBPS were independently proposed by Goetschalckx et al.~\cite{goetschalckx2011continuous} and Montesano and Lopes~\cite{montesano2012active} to model correlations between the Bernoulli random variables in a Beta-Bernoulli process.
Such correlations may exist when the Bernoulli random variables depend on common latent factors.
For example, two grasps may have similar $P_F$ when they contact the same shape in similar locations or the surface geometry at the contact points is similar, as evidenced by Lipschitz bounds on grasp WS metrics~\cite{pokorny2013c}.

Let $\mO$ denote the input object to Algorithm~\ref{alg:full}, and let $\Gamma$ be the set of $N_g$ candidate grasps generated for $\mO$.
We define $F_{j} = F(\bg_{j}) \in \{0, 1\}$ as force closure on an evaluation of any grasp $\bg_{j} \in \Gamma$ from samples of object pose, gripper pose, and friction as described in \secref{quality}.
Under our model, $F_j$ is a Bernoulli random variable with probability of success $\theta_{j}= P_F(\bg_{j})$.
Since $\theta_j$ is unknown, our algorithm maintains a posterior belief distribution for each $\theta_j$ based on the outcome of past force closure evaluations.
In CCBPs, the belief distribution on the Bernoulli parameter $\theta_{i,k}$ is a Beta distribution, which is specified by shape parameters $\alpha > 0$ and $\beta > 0$:

\vspace{-2ex}
\begin{align*}
	\betadist(\alpha, \beta) = \frac{1}{B(\alpha, \beta)} \theta_k^{\alpha-1} (1 - \theta_k)^{\beta-1}
\end{align*}

A CCBP estimates the shape parameters for a grasp-object pair $\mY_{j} = (\bg_{j}, \mO_{i}) \in \mM$ using a normalized kernel function $k(\mY_p, \mY_q) : \mM \times \mM \rightarrow [0,1]$ that measures similarity between a pair of grasps and objects from the Grasp Moduli Space $\mM$.
The kernel approaches 1 as the arguments become increasingly similar and approaches 0 as the arguments become dissimilar.
We measure similarity as a function of the euclidean distance between the output of a set of $N_f$ feature maps $\phi_m: \mM \rightarrow \mathbb{R}^{d_m}$ for $m = 1, ..., N_f$, where $d_m$ is the dimension of the feature space.
Details on our feature maps are given in \secref{features} below.
Given the feature maps, we use the squared exponential kernel 
\begin{align*}
	k(\mY_p, \mY_q) &= \exp\left( - \frac{1}{2} \sum \limits_{m=1}^{N_f} \|\phi_m(\mY_p) - \phi_m(\mY_q)\|_{C_m}^2 \right).
\end{align*}
\noindent where $C_m \in \bR^{d_m \times d_m}$ is the bandwidth for $\phi_m$ and $\| \by \|_{C_m} = \by^T (C_m^T C_m)^{-1} \by$.
The bandwidths are hyperparameters of the algorithm that might be set by cross-validation on a set of training data~\cite{rasmussen2006}.

We form a belief distribution for each candidate grasp in $\Gamma$ based on its its similarity to all grasps and objects from the Dex-Net 1.0 database $\mD$ as measured by the kernel~\cite{goetschalckx2011continuous}:
%\vspace{-2ex}
\begin{align}
	%p \left(\theta_{j} | \mD \right) &= \betadist\left( \alpha_{j,0}, \beta_{j,0} \right) \notag \\
	\alpha_{j,0} = \alpha_{0} & + \sum \limits_{i=1}^{N_o} \sum \limits_{k=1}^{N_g} k(\mY_{j}, \mY_{i, k}) S_{i,k} \label{eq:alpha-prior} \\
	\beta_{j,0} = \beta_{0} & + \sum \limits_{i=1}^{N_o} \sum \limits_{k=1}^{N_g}  k(\mY_{j}, \mY_{i,k}) (N_s - S_{i,k}) \label{eq:beta-prior}
\end{align}
\noindent where $\alpha_{0}$ and $\beta_{0}$ are prior parameters for the Beta distribution~\cite{laskey2015bandits} and $N_s$ is the number of times each grasp in $\mD$ was sampled for evaluate $P_F$.
In practice, we estimate the above sums using the $N_n$ nearest neighbors to $\mO$ in the object similarity KD-Tree descirbed in \secref{object-similarity}.
Upon observing $F_{\ell}$ for grasp $\bg_{\ell}$ on iteration $t$, we update our belief for all other grasps on object $\mO$ by~\cite{goetschalckx2011continuous}:

\vspace{-2ex}
\begin{align}
	\alpha_{j,t} &= \alpha_{j,t-1} + k(\mY_{j}, \mY_{\ell}) F_{\ell} \label{eq:alpha} \\
	\beta_{j,t} &= \beta_{j,t-1} + k(\mY_{j}, \mY_{\ell}) (1 - F_{\ell})\label{eq:beta}
\end{align}

\noindent Intuitively, this allows evaluations of a grasp on an object to constitute fractional evaluations of similar grasps on similar objects.

\subsection{Feature Representation}
\seclabel{features}
\TODO{Condense section / possibly move earlier.}
We feature maps based on the grasp parameters $\phi_g$, the object local surface geometry $\phi_h$, and the global object shape $\phi_s$ to capture correlations between the $P_F$ of grasps.

\subsubsection{Grasp Parameters}
\seclabel{param-features}
Our first feature map is $\phi_{g}(\mY) = (\bx, \bv, \| \rho_1 \|_2, \| \rho_2 \|_2)$ where $\bx \in bR^3$ is the grasp center, $\bv \in \bS^2$ is the grasp approach, and $\rho_i \in \bR^3$ is the $i$-th moment arm.
The moment arms are included because points on the object surface with larger moment arms will move greater distances under object orientation uncertainty.
However, similarity $\phi_g$ may not indicate similar $P_F$ if the surfaces near the contact locations are dissimilar (consider two grasps that contact a box near the corner).

\subsubsection{Local Surface Heightmaps}
\seclabel{heightmaps}
We also use a variant of the grasp heightmap features of Herzog et al.~\cite{herzog2012template} and Kappler et al.~\cite{kappler2015leveraging} to capture local surface geometry.
Let $d_h \in \mathbb{Z}$ be the number of pixels along each dimension of the heightmap, let $\mP = \{-d_h, ..., d_h\}$ be the row / column pixel indices for the heightmap, let $\delta \in \mathbb{R}$ be the resolution of the image pixels in meters, and let $r \in \mathbb{R}$ be a minimum projection distance.
Furthermore, let $\bc_i, i = 1, 2$ be a contact point for grasp $\bg$ and let $\bt_1, \bt_2$ be two orthogonal unit vectors to the grasp approach direction $\bv$.
Our heightmap at contact $i$, $\bh_{i}: \mP \times \mP \rightarrow \bR$, maps discrete locations along the tangent plane specified by $\bt_1, \bt_2$ to the distance to the surface along the grasp axis $\bv$.
To compute the heightmap value at pixel $u,v \in \mP$, we first compute the 3D location of the pixel on the plane $\bp_i(u,v) = \bc_i + \delta u \bt_1 + \delta v \bt_2$.
Then we assign the heightmap value
\begin{align*}
	\bh_i(u,v) = \minimum{t \geq -r} t \text{ such that } f\left( \bp_i(u,v) + s_i \bv \right) = 0
\end{align*}
\noindent where $s_1 = -1$, $s_2 = 1$, and $f$ is the SDF of object $\mO$. 
We then make $\bh_i$ rotation-invariant by orienting its axes to align with the eigenvectors of the weighted covariance matrix of the 3D surface points that generate the heightmap as described in~\cite{tombariunique}.
\figref{local-feature-model} illustrates local surface patches extracted by this procedure.
Since force closure depends largely on object surface normals at the contact points, we finally take the $x$- and $y$-image gradients to form local surface normal maps $\bd_{i,x}$ and $\bd_{i,y}$.
Our full feature vector for the heightmaps is $\phi_{h}(\bg_i, \mO_i) = (\bd_{1,x}, \bd_{1,y}, \bd_{2,x}, \bd_{2,y})$.

\begin{figure}[t!]
\centering
\includegraphics[scale=0.35]{figures/illustrations/local_feature_model.png}
\caption{Three local surface heightmaps extracted on a teapot. Each heightmap is ``rendered" along the grasp axis at each contact point and oriented by the local directions of maximum variation in the heightmap.  We extract the gradients of these heightmaps as features in our robust grasp planning algorithm.}
\figlabel{local-feature-model}
\vspace*{-15pt}
\end{figure}

\subsubsection{Shape Features}
\seclabel{global-sim}
While force closure is based on local surface properties of an object around contact points, measuring global similarity may be useful because the heightmap gradients may not capture all possible contact locations for a grasp under perturbations in object and gripper pose.
Thus we also use the feature map $\phi_{s}(\mY) = \psi(\mO)$, where $\psi$ is the similarity embedding based on MV-CNNs described in \secref{object-similarity}.

%\subsection{Thompson Sampling}
%\seclabel{thompson}
%
%Following Laskey et al.~\cite{laskey2015bandits}, we use Thompson sampling to select the next grasp to evaluate given belief distributions on the $P_F$ for each grasp on iteration $t$ specified by estimates $\mA_t = \{\alpha_{1,t}, ..., \alpha_{K,t}\}$ and $\mB_t = \{\beta_{1,t}, ..., \beta_{K,t}\}$.
%Thompson sampling samples a probability of force closure $\hat{\theta}_{j,t} \sim B(\alpha_{j,t}, \beta_{j,t})$ for all grasps $j = 1, ..., K$, then selects the grasp $j_t^* = \myargmax{j} \hat{\theta}_{j,t}$ to evaluate next.
%Several other criteria exist for selecting the next evaluation in MAB, such as Gittins indices~\cite{laskey2015bandits}, Upper Confidence Bounds~\cite{boularias2015learning}, and Expected Improvement~\cite{montesano2012active}.