### 2015-09-10 ICRA2016/bandits_docs/includes/conclusion.tex

\section{Discussion and Future Work}
\seclabel{conclusion}
We developed a Multi-Armed Bandit algorithm based on correlated rewards to reduce the number of samples required to plan robust parallel-jaw grasps by leveraging prior grasps and 3D object models.
To study scaling effects, we introduced the Dexterity Network (Dex-Net) 1.0, a dataset of approximately 10,000 3D object models, each labelled with up to 250 grasps and the probability of force closure of each.
For efficient indexing of similar objects in Dex-Net, we used multi-view Convolutional Neural Networks (MV-CNNs)~\cite{su2015multi}, a state-of-the-art deep learning method of object classification, trained on milliions of images and thousands of 3D object examples.
Our experiments suggest that increasing the amount of prior knowledge of objects and the quality of grasps on each object enables up to a $10\times$ speedup in the time to compute a grasp with high probability of force closure for a new object, and also suggest that convergence is fastest when the object has a set of geometrically similar nearest neighbors in the database.

One current shortcoming of our method is that it relies on a set of features manually designed to predict grasp quality for a grasp and object, such as heightmaps and CNNs that predict object category.
Future work will leverage recent developments in feature learning with Deep Neural Networks~\cite{krizhevsky2012imagenet} to optimize grasp and object representations directly based on the outcome of grasp evaluations from bandit algorithms.
We are also interested in a future pipelined approach that integrates research on deep learning to predict object identity and pose directly from depth images as input to Dex-Net~\cite{guptaCVPR15a} and deep learning for control policies to map planned grasps on objects to motor torques~\cite{levine2015end}. 
We will use supervision from the data in Dex-Net to study extensions of our MV-CNNs to predict 3D object model and pose from simulated depth images of objects on a planar work surface to research deep learning approaches that map from 3D objects directly to grasp or manipulation policies, with the goal of a future system that maps directly from images to motor torques.

Future work will also add additional grasp and object data to Dex-Net, such as task constraints, the outcomes of physical trials, or human annotations.
%This presents a challenge because the number of labels per object and grasp may be smaller that with probability of force closure due to the time cost involved.
%Thus we will study using functional map networks to establish consistent maps between objects to share sparse labels throughout Dex-Net~\cite{huang2013fine}.
We plan to add additional analytic grasp metrics to Dex-Net, such as cages~\cite{diankov2010automated} or the expected Ferrari-Canny quality~\cite{kim2012physically}, and study the selection of a set of grasps that ``cover" an object surface to increase the chances that a grasp is reachable in cluttered environments.
We plan to release Dex-Net in the future as an open-access project with an open-source API for labelling objects in the Cloud with analytic quality metrics or simulation outcomes and integrating Dex-Net with physical robots.