### 2015-09-14 ICRA2016/bandits_docs/includes/abstract.tex
Long version of the paper is now near-finished. From this point on I will work on trimming

\begin{abstract}
Recent advances in distributed computing and machine learning combined with very large datasets of images and human speech have produced a series of results that surpass decades of research in analytic methods and, in some cases, human ability.
This raises the question: can Big Data and Cloud computation produce analogous advances in robot grasping and manipulation?
We develop a Multi-Armed Bandit algorithm with correlated rewards and show that our algorithm can leverage a large dataset of prior grasps and 3D object models to reduce the number of samples needed for robust grasp planning.
To study scaling effects, we introduce the Dexterity Network (Dex-Net) 1.0, a growing dataset that currently includes over 10,000 unique 3D object models and 2.5 million parallel-jaw grasps along with the probability of force closure of each under uncertainty in object pose, gripper pose, and friction coefficient.
We use multi-view Convolutional Neural Networks (MV-CNNs), a new deep learning method for 3D object classification, to learn a similarity metric between objects and apply MV-CNNs to indexing similar objects for grasp planning at scale.
We implement Dex-Net 1.0 using the Google Cloud Platform with a system that can run up to 1,500 virtual machines at a time, reducing experiment runtime by up to 1000$\times$.
Experiments suggest that robust grasp planning is up to $3.5\times$ faster than Thompson sampling without prior data on average and that average grasp quality scales with the size of the dataset when the MV-CNN returns geometrically similar objects.%, and that the convergence rate can descrease if the similarity metric measures false correlations between grasps.
We also explore sensitivity of our algorithm to similarity metrics and pose and friction uncertainty levels for test objects.
Code, data, and additional information can be found at: \TODO{Url}
\end{abstract}