### 2015-09-15 ICRA2016/bandits_docs/includes/problem_statement.tex

\section{Definitions and Problem Statement}
\seclabel{problem}

We consider the robust grasp planning problem for a single 3D object model and parallel-jaw grippers using probability force closure, $P_F$,  under uncertainty in object pose, gripper pose, and friction coefficient as a grasp quality metric.
We assume the object shape is provided as a signed distance function (SDF) $f: \mathbb{R}^3 \rightarrow \mathbb{R}$~\cite{mahler2015gp}, which is zero on the object surface, positive outside the object, and negative within. 
The object is specified in units of meters with given center of mass $\bz \in \mathbb{R}^3$.
We assume soft-finger contacts with a Coulomb friction model~\cite{zheng2005} and that the gripper jaws are opened to their maximal width $w \in\bR$ before closing on the object.
%and known uncertainty in object pose, gripper pose relative to the object, and friction coefficient, described in detail in \seclabel{uncertainty} below.

\subsection{Grasp and Object Parameterization}
\seclabel{grasp-param}
Our grasp parameterization is illustrated in~\figref{grasp-model}.
Let $\bg = (\bx, \bv)$ be a parallel-jaw grasp parameterized by the centroid of the jaws in 3D space $\bx \in \mathbb{R}^3$ and an approach direction, or axis, $\bv \in \bS^2$.
%Let $f: \mathbb{R}^3 \rightarrow \mathbb{R}$ be a signed distance function (SDF) representing an object's geometry~\cite{mahler2015gp, newcombe2011kinectfusion}.
We denote by $\mS = \{\by \in \bR^3 \big| f(\by) = 0\}$ the surface of an object for SDF $f$, and specify all points with respect to a reference frame centered at the object center of mass $\bz$ and oriented along the principal axes of $\mS$.
Let $\mG = \{ (\bx, \bv) \big| \bx \in \mathbb{R}^3, \bv \in \bS^2\}$ denote the space of all grasps and $\mH = \{ \mO = \{\bz, f(\cdot)\} \big|  \bz \in \mathbb{R}^3, f \in \mA\}$ denote the space of all objects, where $\mA$ is the space of all SDFs for closed and compact surfaces.
We denote by $\mM = \mG \times \mH$ the joint space of all parallel-jaw grasps and objects, also known as a Grasp Moduli Space~\cite{pokorny2013grasp}.

\begin{figure}[t!]
\centering
\includegraphics[scale=0.44]{figures/illustrations/dexnet_grasping_model.eps}
\caption{Illustration of grasp parameterization and contact model. (Left) We parameterize parallel-jaw grasps by the centroid of the jaws $\bx \in \mathbb{R}^3$ and approach direction, or direction along which the jaws close, $\bv \in \bS^2$. The parameters $\bx$ and $\bv$ are specified with respect to a coordinate frame at the object center of mass $\bz$ and oriented along the principal directions of the object. (Right) The jaws are closed until contacting the object surface at locations $\bc_1, \bc_2 \in \mathbb{R}^3$, at which the surface has normals $\bn_1, \bn_2 \in \bS^2$. The contacts are used to compute the moment arms $\rho_i = \bc_i - \bz$.
% and $\rho_2  = \bc_2 - \bz$.
}
\figlabel{grasp-model}
\vspace*{-15pt}
\end{figure}

\subsection{Sources of Uncertainty}
\seclabel{uncertainty}
We assume Gaussian distributions on object pose, gripper pose, and friction coefficient.
Let $\upsilon \sim \mN\left(\mathbf{0}, \Sigma_{\upsilon}\right)$ denote a zero-mean Gaussian on $\bR^6$ and $\mu_{\xi} \in SE(3)$ be a mean object pose.
We define the object pose random variable $\xi = \exp\left( \upsilon^{\wedge} \right) \mu_{\xi}$, where the $\wedge$ operator maps from $\bR^6$ to the Lie algebra $\mathfrak{se}(3)$.
Let $\nu \sim \mN\left(\mu_{\nu}, \Sigma_{\nu}\right)$ denote a Gaussian distribution on gripper pose with mean $\mu_{\nu} \in \mG$.
Let $\gamma \sim \mN\left(\mu_{\gamma}, \Sigma_{\gamma} \right)$ denote a Gaussian distribution on friction coefficient with mean $\mu_{\gamma} \in \mathbb{R}$.
We denote by $\hat{\xi}, \hat{\nu},$ and $\hat{\gamma}$ samples of object pose, gripper pose, and friction coefficient, respectively. 
%Theses sources of uncertainty might be estimated based on prior models of repeatability in gripper actuation or from the posterior variance in object pose or friction coefficient in estimates based on sensor data.

\subsection{Contact Model}
\seclabel{contact}
%Our contact model is illustrated in the right panel .
Given a grasp $\bg$ on an object $\mO$ and samples $\hat{\xi}, \hat{\nu},$ and $\hat{\gamma}$, let $\bc_i \in \mathbb{R}^3$ for $i \in {1, 2}$ denote the 3D contact location between the $i$-th gripper jaw and surface as shown in \figref{grasp-model}.
Each contact $\bc_i = \bx + (-1)^{i} (t_i^* - w / 2) \bv$ where~\cite{mahler2015gp}:
%\vspace{-1ex}
\begin{align*}
	t_i^* &= \minimum{t \geq 0} t \text{ such that } f\left(\bx +  (-1)^{i} (t - w / 2) \bv\right) = 0.
\end{align*}
\noindent Let $\bn_i = \nabla f(\bc_i) / \| \nabla f(\bc_i) \|_2$ denote the surface normal at contact $\bc_i$ with tangent vectors $\bt_{i,1}, \bt_{i,2} \in \bS^2$.
To compute the forces that each contact can apply to the object for friction coefficient $\hat{\gamma}$, we discretize the friction cone at $\bc_i$~\cite{pokorny2013c} into a set of $l$ facets with vertices $\mF_{i} = \left\{ \bn_i + \hat{\gamma} \cos \left( \frac{2 \pi j}{l} \right) \bt_{i,1} + \hat{\gamma} \sin \left( \frac{2 \pi j}{l} \right) \bt_{i,2} \big| j = 1, ..., l \right\}$.
Each force $\bbf_{i,j} \in \mF_{i}$ can exert a corresponding torque $\tau_{i,j} = \bbf_{i,j} \times \rho_i$ where $\rho_i = (\bc_i - \bz)$ is the moment arm at $\bc_i$.
Under our soft contact model, each contact $\bc_i$ exerts an additional wrench $\bw_{i,l+1} = (\mathbf{0}, \bn_i)$~\cite{zheng2005}.
Thus the set of all contact wrenches that can be applied by a grasp $\bg$ under our model is $\mW = \{ \bw_{i,j} = (\bbf_{i,j}, \tau_{i,j}) \big| i = 1, 2 \text{ and } j = 1, ..., l+1\}$.

\subsection{Quality Metric}
\seclabel{quality}
In this work we use the probability of force closure ($P_F$), or the ability to resist external force and torques in arbitrary directions~\cite{mahler2015gp}, as our quality metric.
$P_F$ has shown promise in physical experiments~\cite{kim2012physically, weisz2012pose} and is relatively inexpensive to evaluate, allowing us to better study the effects of large amounts of data.

Let $F \in \{0, 1\}$ denote the occurrence of force closure.
Then for a grasp $\bg$ on object $\mO$, $P_F(\bg, \mO) = \mathbb{P}\left(F = 1 \mid \bg, \mO, \xi, \nu, \gamma \right)$.
To compute force closure for a grasp $\bg \in \mG$ on object $\mO \in \mH$ given samples of object pose $\hat{\xi}$, gripper pose $\hat{\nu}$, and friction coefficient $\hat{\gamma}$, we first compute the set of possible contact wrenches $\mW$.
Then $F = 1$ if the origin lies within the convex hull of the contact wrench set, $\mathbf{0} \in Conv(\mW)$~\cite{weisz2012pose}.

\subsection{Objective}
We are interested in finding $\bg^*$, the grasp that maximizes $P_F(\bg)$~\cite{kim2012physically, laskey2015bandits, mahler2015gp, weisz2012pose} over a budgeted maximum number of samples $T$.
To perform this as quickly as possible we maximize over the sum of the $P_F$ for all sampled grasps~\cite{laskey2015bandits, srinivas10gaussian}.
As the maximization over the continuous space $\mG$ is computationally expensive, past work has solved this objective by evaluating a discrete set of $K$ candidate grasps $\Gamma = \left\{ \bg_1, ..., \bg_K \right\}$ with Monte-Carlo integration~\cite{kehoe2012toward, weisz2012pose} or Multi-Armed Bandits (MAB) ~\cite{laskey2015bandits}.
In this work, we extend the MAB model of~\cite{laskey2015bandits} to 3D and to leverage similarities between grasps and prior objects in Dex-Net to reduce the samples needed to converge to $\bg^*$.