### 2015-09-15 ICRA2016/bandits_docs/includes/conclusion.tex

\section{Discussion and Future Work}
\seclabel{conclusion}
We presented Dexterity Network 1.0 (Dex-Net), a new dataset and associated algorithm to study the scaling effects of Big Data and Cloud Computation on robust grasp planning.
The algorithm uses a Multi-Armed Bandit model with correlated rewards to leverage prior grasps and 3D object models and Multi-View Convolutional Neural Networks (MV-CNNs), a new deep learning method for 3D object classification, as a similarity metric between objects.
In experiments, the Google Cloud Platform allowed Dex-Net 1.0 to simultaneously run up to 1,500 virtual machines, reducing experiment runtime by three orders of magnitude.
Experiments suggest that prior data can speed robust grasp planning by a factor of 3 and that average grasp quality increases with the number of similar objects in the dataset.
We reported on system sensitivity to varying similarity metrics and pose and friction uncertainty levels.

In future work, we will develop metrics to pre-compute grasps that adequately "cover" each object from a variety of accessibility conditions (depending on pose and occlusions).
We will also explore how Deep Learning~\cite{krizhevsky2012imagenet} can be used in other parts of a grasp planning pipeline, for example to recognize object pose and shape from images~\cite{aubry2015understanding}, to learn grasp and object features robust to shape variation using prior evaluations from bandit algorithms, and perhaps even to determine motor torques based on images and precomputed grasps~\cite{levine2015end}.
We also hope to release subsets of Dex-Net 1.0 with an open-source API to explore robust grasping as a service (RGaaS).

%We developed the Dex-Net 1.0 algorithm using a Multi-Armed Bandit model with correlated rewards to reduce the number of samples required to plan robust parallel-jaw grasps by leveraging prior grasps and 3D object models.
%To study scaling effects, we introduced the Dexterity Network (Dex-Net) 1.0, a dataset of approximately 10,000 3D object models and 2.5 million parallel-jaw grasps, and MV-CNNs~\cite{aubry2015understanding, su2015multi} to efficiently index similar objects.
%Our experiments suggest that increasing the amount of prior knowledge of objects and the quality of grasps on each object enables up to a $10\times$ speedup in the time to compute a grasp with high probability of force closure for a new object, and also suggest that convergence is fastest when the object has a set of geometrically similar nearest neighbors in the database.
%
%One current shortcoming of our method is that it relies on a set of features manually designed to predict grasp quality for a grasp and object, such as heightmaps and CNNs that predict object category.
%Future work will leverage recent developments in feature learning with Deep Neural Networks~\cite{krizhevsky2012imagenet} to optimize grasp and object representations directly based on the outcome of grasp evaluations from bandit algorithms.
%We are also interested in an end-to-end approach that integrates research on deep learning to map directly from images to motor torques that control a grasp.
%To do so, we will use supervision from the data in Dex-Net to study prediction of object identity and pose directly from depth images~\cite{aubry2015understanding} and deep learning for control policies to map planned grasps on objects to motor torques~\cite{levine2015end}. 
%%We will use supervision from the data in Dex-Net to study extensions of our MV-CNNs to predict 3D object model and pose from simulated depth images of objects on a planar work surface and to research deep learning approaches that map from 3D objects directly to grasp or manipulation policies, with the goal of a future system that maps directly from images to motor torques.
%
%In future work, we will also add grasp and object data to Dex-Net, such as more 3D models, task constraints, the outcomes of physical trials, or human annotations.
%%This presents a challenge because the number of labels per object and grasp may be smaller that with probability of force closure due to the time cost involved.
%%Thus we will study using functional map networks to establish consistent maps between objects to share sparse labels throughout Dex-Net~\cite{huang2013fine}.
%We will also explore additional analytic grasp metrics to Dex-Net such as cages, different grasp selection policies such as Gittins indices, and extensions to plan a set of grasps that ``cover" an object surface to increase the chances that a grasp is reachable in cluttered environments.
%We will also include uncertainty in object shape resulting from noisy point clouds~\cite{mahler2015gp} to future models.
%Our goal is to release Dex-Net in the future as an open-access project with an open-source API for labelling objects in the Cloud with analytic quality metrics or simulation outcomes and integrating Dex-Net with physical robots.
%